Effects without a Cause: The Search for Demographic Anomalies

### 2020-01-20 MANUSCRIPT/manuscript.Rmd
Update manuscript.Rmd
Some changes to results section

**\autoref{sumtable}** report the overall number of anomalies we detect of each type for births and deaths and some summary statistics across all anomaly types. We find considerably more mortality anomalies (n = `r paste(sums[2,2])`) than fertility anomalies (n = `r paste(sums[1,2])`). This is likely because mortality is more easily subject to instantaneous change compared with fertility. Changes in fertility should theoretically manifest approximately 9-months after change.

We also find more anomalous deaths (`r paste(prettyNum(sums[2,6], digits = 0, big.mark = ","))`) than anomalous births (`r paste(prettyNum(sums[1,6], digits = 0, big.mark = ","))`). Fertility anomalies overwhelmingly tended toward lower fertility, with only a few fertility anomalies yielding more births. Conversely, mortality anomalies tended to be more evenly split between positive and negative anomalies, but tended toward more deaths rather than less deaths. These anomalous deaths and births account for `r sums2$percent_anom[2]`% and `r sums2$percent_anom[1]`% over the entire time series, respectively.

>>>>

**\autoref{sumtable}** report the overall number of anomalies we detect of each type for births and deaths and some summary statistics across all anomaly types. We find considerably more mortality anomalies (n = `r paste(sums[2,2])`) than fertility anomalies (n = `r paste(sums[1,2])`). Given that anomalies are always related to events [@Song2018ACM], the finding of more mortality than fertility anomalies is not surprising.  Mortality is more easily subject to instantaneous change (like a natural disaster) compared with fertility, and fertility is linked to human decision-making more strongly than mortality [@Stein2014Fertility]. Previous studies document that some events, like the 2005 Oklahoma City bombings, resulted in both fertility and mortality changes [@Rogers2005OKBombing], but changes in fertility manifest approximately 9-months after an event, and the relationship between a fertility-inducing event and behavior change (such as a prolonged blizzard or power blackout) is less strong than the relationship between a mortality-inducing event (such as an earthquake) and death. For example, Evans and his colleagues [@Evans2008Hurricanebirth] show that catastrophic weather events can lead to increased fertility, but some events reduce fertility.  Furthermore, the effect is seen primarily in those who already have one child, with childless people unlikely to have a child due to a weather event.

We also find more anomalous deaths (`r paste(prettyNum(sums[2,6], digits = 0, big.mark = ","))`) than anomalous births (`r paste(prettyNum(sums[1,6], digits = 0, big.mark = ","))`). Fertility anomalies overwhelmingly tend toward lower fertility, with only a few fertility anomalies yielding more births. This is consistent with research findings that link high severity weather events to significantly lower fertility [@Evans2008Hurricanebirth]. Conversely, mortality anomalies tend to be more evenly split between positive and negative anomalies, but tend toward more deaths rather than less deaths. Again, this is not surprising, as the kinds of events that increase death (e.g., plant closings are associated with increased opioid death [@Venkataramani2019]) are more common than the events that decrease death (e.g., cannabis legalization is associated with decreased opioid deaths [@Livingston2017Cannabis]) These anomalous deaths and births account for `r sums2$percent_anom[2]`% and `r sums2$percent_anom[1]`% over the entire time series, respectively.
